---
title: "Homework 6"
author: "Alyssa Vanderbeek"
date: "27 November 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

theme_set(theme_bw())
```

### Problem 1

```{r}
homicide_data = read_csv('./data/problem1/homicide-data.csv') # read in data

homicide_data_tidy = homicide_data %>%
  mutate(city_state = paste(city, toupper(state), sep = ', '), # create city_state variable
         victim_race = fct_relevel(ifelse(victim_race == 'White', 'white', 'non-white'), 'white'), # dichotomize race as factor (white vs. non-white)
         victim_age = ifelse(victim_age == 'Unknown', NA, as.numeric(victim_age)), # make age numeric
         resolved = as.numeric(disposition == "Closed by arrest")) %>% # create binary variable for whether murder is resolved
  filter(!(city_state %in% c('Tulsa, AL', 'Dallas, TX', 'Phoenix, AZ', 'Kansas City, MO'))) # filter cities

str(homicide_data_tidy)

# Logistic regression for all cities
fit_logistic = homicide_data_tidy %>%
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())

summary(fit_logistic)
  
# Table of estimated OR and 95% CI for race, holding all else constant
fit_logistic %>% 
  broom::tidy() %>% # tidy output
  filter(term == 'victim_racenon-white') %>% # get coefficient for race only
  summarise(OR = exp(estimate),
            '95% CI lower bound' = exp(estimate - qnorm(0.975)*std.error),
            '95% CI upper bound' = exp(estimate + qnorm(0.975)*std.error)) %>% # calculate 95% CI for OR
  knitr::kable(digits = 3)
```

Across all cities in the dataset, the odds of the murder of a non-white person being resolved is 44% less than if the victim were white (all else constant). 

```{r}
# Logistic regression for each city
log_reg_cities = homicide_data_tidy %>%
  select(city_state, resolved, victim_age, victim_race, victim_sex) %>% # select only variables to use in logistic regression
  nest(-city_state) %>% # nest all variables to each city
  mutate(logreg = map(data, ~ broom::tidy(glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())))) %>% # map logistic regression with nested data, make output a nested tibble
  select(city_state, logreg) %>% # select only city and model
  unnest %>% 
  filter(term == 'victim_racenon-white') %>% # get coefficients for race only
  rowwise %>%
  summarise(city = city_state,
            OR = exp(estimate),
            ci_lower = exp(estimate - qnorm(0.975)*std.error),
            ci_upper = exp(estimate + qnorm(0.975)*std.error)) # OR and 95% CI for each city

# print table
log_reg_cities %>%
  knitr::kable(digits = 3,
               col.names = c('City', 'OR', '95% CI lower bound', '95% CI upper bound'),
               caption = 'Odds of resolved cases for non-white vs. white victims') # write an actual caption
```


```{r, fig.height=7, fig.width=5}
# plot of OR estimate and CI for each city
log_reg_cities %>%
  mutate(city = fct_reorder(city, OR, desc = T)) %>% # order cities according to OR estimate
  ggplot(aes(x = city, y = OR)) + 
  geom_point() +
  geom_errorbar(aes(x = city, ymin = ci_lower, ymax = ci_upper), width = 0.5) + # error bars for CI bounds
  geom_hline(yintercept = 1, lty = 3) + # dotted line at OR=1 to represent no difference in odds of resolution across races
  labs(y = 'Odds of resolved cases (non-white vs. white)',
       x = 'City') +
  coord_flip()
  
```

The figure above shows the odds ratios for case resolution of white vs. non-white victims (all else constant). All cities on display expect for two - Tampa, FL and Birmingham, AL - exhibit a lower odds of resolving murder of non-white victims, compared to white victims. The odds of resolution between races is about equal in Durham, NC, but the confidence interval around the estimate is wide. 

### Problem 2

```{r}
birthweight = read_csv('./data/problem2/birthweight.csv')

birthweight_tidy = birthweight %>%
  mutate(babysex = fct_recode(as.character(babysex), 'Male' = '1', 'Female' = '2'),
         mrace = fct_recode(as.character(mrace), 'White' = '1', 'Black' = '2', 'Asian' = '3', 'Puerto Rican' = '4'),
         frace = fct_recode(as.character(frace), 'White' = '1', 'Black' = '2', 'Asian' = '3', 'Puerto Rican' = '4', 'Other' = '8'))

str(birthweight)


lm(bwt ~ blength + gaweeks, data = birthweight_tidy) # comparison model 1
lm(bwt ~ (bhead + blength + babysex)^3, data = birthweight_tidy) # comparison model 2 with all interaction terms 

```

